On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-08-15 manuscripts/PNAS/chazthm_pnas.tex

Before stating our main result, we identify combinatorial criteria on the support sets of the generating codes $\mathbf{x}_i$ that imply stable sparse representations. Letting $\{1, \ldots, m\}$ be denoted $[m]$, its power set $2^{[m]}$, and ${[m] \choose k}$ the set of subsets of $[m]$ of size $k$, we say a hypergraph $\mathcal{H} \subseteq 2^{[m]}$ on vertices $[m]$ is \textit{$k$-uniform} when in fact $\mathcal{H} \subseteq {[m] \choose k}$. The \emph{degree} $\deg(i)$ of a node $i \in [m]$ is the number of sets in $\mathcal{H}$ that contain $i$, and we say $\mathcal{H}$ is \emph{regular} when for some $r$ we have $\deg(i) = r$ for all $i$ (for given $r$, we say $\mathcal{H}$ is \textit{$r$-regular}). We also write $2 \mathcal{H} := \{ S \cup S': S, S' \in \mathcal{H}\}$.

@@

[*** We remark about algorithms ***]

We remark that there is a strong application to theoretical neuroscience.

\begin{corollary}
Unsupervised neurally plausible sparse feature learning through a wiring bottleneck recovers original sensory representations and neural activities.
\end{corollary}

[*** Make clearer / motivate better ***]

>>>>

Before stating our main result, we identify combinatorial criteria on the support sets of the generating codes $\mathbf{x}_i$ that imply stable sparse representations. Letting $\{1, \ldots, m\}$ be denoted $[m]$, its power set $2^{[m]}$, and ${[m] \choose k}$ the set of subsets of $[m]$ of size $k$, we say a hypergraph $\mathcal{H} \subseteq 2^{[m]}$ on vertices $[m]$ is \textit{$k$-uniform} when in fact $\mathcal{H} \subseteq {[m] \choose k}$. The \emph{degree} $\deg(i)$ of a node $i \in [m]$ is the number of sets in $\mathcal{H}$ that contain $i$, and we say $\mathcal{H}$ is \emph{regular} when for some $r$ we have $\deg(i) = r$ for all $i$ (for given $r$, we say $\mathcal{H}$ is \textit{$r$-regular}). We also write $2\mathcal{H} := \{ S \cup S': S, S' \in \mathcal{H}\}$.

@@

We remark that these results have a strong application to theoretical neuroscience, justifying what may be the first hypothesized theory of compressive communication between sparsely active neural populations \cite{Isely10, ganguli2012compressed}.

\begin{corollary}
Random noisy (linear) compression of sparse neural activities $\mathbf{x}_i$ is decompressible by any neurally plausible unsupervised sparse feature learning algorithm.
\end{corollary}
